---
title: "Covid-19 Update"
author: "Tommaso Manara"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show 
---


```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)

Sys.setlocale("LC_TIME", "English")
```


```{r load-libraries, include=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)
```


#CDC COVID-19 Public Use Data

Let us revisit the [CDC Covid-19 Case Surveillance Data](https://data.cdc.gov/Case-Surveillance/COVID-19-Case-Surveillance-Public-Use-Data/vbim-akqf). There are well over 3 million entries of individual, de-identified patient data. Since this is a large file, I suggest use `vroom` to load it and keep `cache=TRUE` in the chunk options.


```{r, cache=TRUE,warnings= FALSE, message=FALSE}
# file contains 11 variables and 3.66m rows and is well over 380Mb. 
# It will take time to download

# URL link to CDC to download data
url <- "https://data.cdc.gov/api/views/vbim-akqf/rows.csv?accessType=DOWNLOAD"

covid_data <- vroom::vroom(url)%>% # If vroom::vroom(url) doesn't work, use read_csv(url)
clean_names()

cleaned_data <- covid_data %>% 
  filter(death_yn %in% c("Yes", "No")) %>% #to remove non-meaningful value in death_yn.
  filter(medcond_yn %in% c("Yes","No") & icu_yn %in% c("Yes", "No")) %>% #to remove non-meaningful value in medcond_yn and icu_yn.
  filter(age_group!="Unknown") %>%  #to remove non-meaningful value in age_group.
  filter(sex %in% c("Male","Female")) %>% 
    mutate(
    medcond_yn=case_when(
      medcond_yn=="Yes" ~ "With comorbidities",
      medcond_yn=="No" ~ "Without comorbidities"
    ),
    icu_yn=case_when(
    icu_yn=="Yes" ~ "Admitted to ICU",
    icu_yn=="No" ~ "No ICU"
    ))


```

Given the data we have, we produce two graphs that show death % rate:

1. by age group, sex, and whether the patient had co-morbidities or not
2. by age group, sex, and whether the patient was admited to Intensive Care Unit (ICU) or not.

```{r two_pictures, warnings= FALSE, message=FALSE}
#by age group, sex, and whether the patient had co-morbidities or not.
cleaned_data %>% 
  group_by(sex, medcond_yn, age_group) %>% 
  summarise(death_num = count(death_yn=="Yes"), total_num =n(), death_rate = death_num/n()) %>% 
  
  ggplot(aes(x=death_rate,y=reorder(age_group, death_rate)))+
    geom_col(fill="#003366")+
    theme_bw() +
    labs (title = "Covid death % by age group, sex and whether patient was admitted to ICU", x="", y="") +
    geom_text(aes(label=round(death_rate*100,1)), hjust=-0.1, size=3)+ #ajust size, position, and decimal for barlabels.
    facet_grid(medcond_yn ~ sex)+ #get the 2x2 combined plot, vertical medcond_yn and horizontal sex.
    scale_x_continuous(limits=c(0,0.85),breaks=seq(0,0.8,0.2),label=scales::percent)+ #make all labels in the plot and  axis label percent.
    theme(plot.margin = unit(c(0.1,0.1,0.1,0.1),"cm"))


#by age group, sex, and whether the patient was admited to Intensive Care Unit (ICU) or not.
cleaned_data %>% 
  group_by(sex, icu_yn, age_group) %>% 
  summarise(death_num = count(death_yn=="Yes"), total_num =n(), death_rate = death_num/n()) %>% 
  
  ggplot(aes(x=death_rate,y=reorder(age_group, death_rate)))+
    geom_col(fill="#FF9999")+
    theme_bw() +
    labs (title = "Covid death % by age group, sex and whether patient was admitted to ICU", x="", y="") +
    geom_text(aes(label=round(death_rate*100,1)), hjust=-0.1, size=3)+
    facet_grid(icu_yn ~ sex)+ #get the 2x2 combined plot, vertical icu_yn and horizontal sex.
    scale_x_continuous(limits=c(0,0.85),breaks=seq(0,0.8,0.2),label=scales::percent)+
    theme(plot.margin = unit(c(0.1,0.1,0.1,0.1),"cm"))
```

